---
title: "Washington Crime Analysis Project"
author: "Tyler Hart"
date: "2022-02-14"
output:
  html_document:
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
margin: 0.5in
classoption: landscape
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
- \floatplacement{figure}{H}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(forecast)
library(tseries)
library(stringr)
library(data.table)
library(scales)
library(lunar)
library(kableExtra)
library(magick)
library(webshot)
library(dygraphs)
library(knitr)
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  results = 'asis',
  fig.align = 'center',
  fig.pos = 'ht',
  dpi = 400,
  fig.path='./figures/',
  dev=c('svg','png'),
  out.width='\\textwidth'
  )
library(kableExtra)
library(lubridate)

source("./text_import.R") # Build Spokane crime data frame
source("./Bellevue_Data_Import.R") # Build Bellevue crime data frame
source("./Tacoma_Data_Import.R") # Build Tacoma crime data frame
#source("./categorize_offenses.R") # Categorize offense names
source("./categorize_violence.R") # Categorize violent / non-violent
source("./lunar_seasons.R") # Lunar phase, seasons

source("./additional_parse.R") # Build Spokane crime data frame

# GGPlot geometry options
point_alpha = 0.5
point_size = 3
geom_line_size = 2

### Figure captions ###
cc_by_sa <- "Licensed CC-BY-SA 4.0, Tyler Hart"

# Spokane
spokane_data_source_str <- paste(paste("Created",Sys.Date(), sep = " "),"Data Source: Spokane PD CompStat", sep = "; ")
excluded_districts_note <- "NOTE: Excluding districts SPA-D"
spokane_caption <- paste(spokane_data_source_str, cc_by_sa, sep = "\n")
spokane_caption_exclude_dist <- paste(paste(spokane_data_source_str, excluded_districts_note, sep = "; "), cc_by_sa, sep = "\n")

# Bellevue
bellevue_data_source_str <- paste(paste("Created",Sys.Date(), sep = " "),"Data Source: Bellevue PD", sep = "; ")
bellevue_caption <- paste(bellevue_data_source_str, cc_by_sa, sep = "\n")

# Tacoma
tacoma_data_source_str <- paste(paste("Created",Sys.Date(), sep = " "),"Data Source: Tacoma PD", sep = "; ")
tacoma_caption <- paste(tacoma_data_source_str, cc_by_sa, sep = "\n")
```

# Introduction

Three large cities in Washington - Bellevue, Spokane, and Tacoma - make regular crime reports available. Spokane releases its data on a weekly basis. This project collects and parses that data, and provides up-to-date analysis. The same data is formatted and submitted annually to the FBI NIBRS program as well. A lot of work is required to prepare the data for submission to NIBRS, but it's more granular and includes detailed information that weekly reports lack. Unfortunately, it takes a year or more for current data to be available in the NIBRS portal. It will most likely be late-2022 or early-2023 before 2021's data is ready. We use the weekly reports to identify trends happening now and provide up-to-date analysis. NIBRS-based analysis is a separate project. The additional detail from years past informs our weekly analysis and helps us frame questions better.

When this project began it was strictly focused on crime in Spokane. It's since expanded to include other cities with similar population size that also make crime data available regularly. Here are the three covered cities and their population based on Census Bureau estimates for 2020:

| City| Estimated 2020 Population|
|----:|---------:|
| Spokane|222,050|
| Tacoma| 219,995|
| Bellevue|148,073|

I'd have liked to include the city of Vancouver, as it has a population between Bellevue and Tacoma, but they don't make regular crime data reports available. Seattle is so much larger than all the other cities that I've excluded its crime analysis from this project.

# Incidents Over Time

We visualize incidents over time by totaling up incidents and plotting them daily. Blue and red lines show average values using generalized additive model (GAM) and linear model (LM), respectively. Then we break down data for time series analysis and identify seasonality and trends.

## Spokane

```{r plot.offenses_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "gam", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  labs(
    x = "", 
    y = "Daily Incidents", 
    title = "Spokane Total Incidents", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

## Tacoma
```{r plot.Tacoma_Offenses_Over_Time, fig.pos="H", fig.width=10, fig.height=7.5}
Tacoma_Daily_Offenses %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "gam", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  labs(
    x = "", 
    y = "Daily Incidents", 
    title = "Tacoma Total Incidents", 
    caption = paste(tacoma_data_source_str, cc_by_sa, sep = "\n")
    )
```

## Bellevue
```{r plot.Bellevue_Offenses_Over_Time, fig.pos="H", fig.width=10, fig.height=7.5}
Bellevue_Daily_Offenses %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "gam", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  labs(
    x = "", 
    y = "Daily Incidents", 
    title = "Bellevue Total Incidents", 
    caption = paste(bellevue_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.crime_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Decompose, plot time series
#ts.crime_stl <- stl(ts.crimes[,2],"periodic", robust = TRUE)
ts.crime_stl <- mstl(ts.crimes[,2], lambda = "auto")
# summary(ts.crime_stl)
# IQR(ts.crimes[,2])
autoplot(ts.crime_stl, main = "Spokane Crime Seasonal Decomposition")
```

```{r plot.crime_seasonally_adjusted, fig.pos="H", fig.width=10, fig.height=7.5}
# Decompose time series
ts.crime_sa <- seasadj(ts.crime_stl) # Adjust for seasonality
autoplot(ts.crime_sa, main = "Offenses, Seasonally Adjusted", xlab = "", ylab = "Daily Offenses")
```

```{r plot.offenses_over_time_violence, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses_violence %>%
  ggplot(aes(date, offenses)) + 
  geom_point(aes(color = violence), size = point_size, alpha = point_alpha) +
  scale_color_manual(values = c("cadetblue3","coral1")) +
  theme(legend.position = c(0.925, 0.925), legend.background = element_rect(size=0.5, linetype="solid", colour ="black")) +
  labs(
    x = "", 
    y = "Offenses", 
    color = "Violence",
    title = "Daily Non-Violent & Violent Offenses", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.offenses_over_time_violence_wrap, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses_violence %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  facet_wrap( ~ violence, ncol = 1, scales = "free") +
  labs(
    x = "", 
    y = "Offenses", 
    color = "Violence",
    title = "Daily Non-Violent & Violent Offenses", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.violent_offenses_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses_violence %>%
  filter(violence == "Violent")  %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  labs(
    x = "", 
    y = "Offenses", 
    title = "Daily Violent Offenses", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.non_violent_offenses_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses_violence %>%
  filter(violence == "Non-Violent")  %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  labs(
    x = "", 
    y = "Offenses", 
    title = "Daily Non-Violent Offenses", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.offenses_over_time_pred, fig.pos="H", fig.width=10, fig.height=7.5}
daily_offenses %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "loess", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", fullrange = TRUE, size = geom_line_size) +
  scale_x_date(limits = as.Date(c("2017-09-03","2028-01-01")), date_breaks = "1 year", date_labels = "%Y") +
  labs(
    x = "", 
    y = "Offenses", 
    title = "Projection, Average Daily Offenses",
    subtitle = "CI 0.95, linear mean projections",
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.monthly_offense_totals, fig.width=10, fig.height=7.5}
summarized_crimes %>%
  ggplot(aes(date, sum.monthly_offenses, label = sum.monthly_offenses)) +
  geom_smooth(size = geom_line_size) +
  geom_point(size = point_size, alpha = point_alpha) +
  geom_text_repel() +
  labs(
  x = "",
  y = "Offenses",
  title = "Total Monthly Offenses",
  caption = spokane_caption
  ) +
  expand_limits(y = 0) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.avg_daily_offenses_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
summarized_crimes %>%
  ggplot(aes(date, avg.daily_offenses, label = round(avg.daily_offenses))) +
  geom_smooth(size = geom_line_size) +
  geom_point(size = point_size, alpha = point_alpha) +
  geom_text_repel() +
  labs(
  x = "",
  y = "Avg. Daily Offenses",
  title = "Average Daily Offenses by Month",
  caption = spokane_caption
  ) +
  expand_limits(y = 0) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.boxplot_avg_daily_by_month, fig.width=10, fig.height=7.5}
df.crimes %>% 
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Offenses", 
    title = "Daily Offenses by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  )
```

```{r plot.monthly_offense_comparisons, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
group_by(year,num.month) %>%
summarize(offenses = n()) %>%
ggplot(aes(num.month, offenses, color = year)) +
geom_point(size = point_size, alpha = point_alpha) +
labs(
  x = "Month",
  y = "Offenses",
  title = "Monthly Offenses",
  color = "Year",
  caption = spokane_caption
  ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.weekly_offense_comparisons, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(date >= "2018-01-01") %>%
group_by(year,num.week) %>%
summarize(offenses = n()) %>%
ggplot(aes(num.week, offenses, color = year)) +
geom_point(size = point_size, alpha = point_alpha) +
labs(
  x = "Week",
  y = "Offenses",
  title = "Weekly Offenses",
  color = "Year",
  caption = spokane_caption
  ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r plot.theft_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(offense == "THEFT") %>%
  group_by(date) %>%
  summarize(offenses = n()) %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    x = "", 
    y = "Thefts", 
    title = "Overall Thefts", 
    caption = spokane_caption
    ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.burglaries_seasonally_adjusted, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(date >= "2018-01-01") %>%
  filter(offense == "BURGLARY") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date

### All data time series ###
ts.crimes <- ts(model_data, start = c(2018,1), end = c(2021,end_date), frequency = 365)

# Decompose time series
ts.crime_stl <- stl(ts.crimes[,2],"periodic") # Decompose time series
ts.crime_sa <- seasadj(ts.crime_stl) # Adjust for seasonality
autoplot(ts.crime_sa, main = "Burglaries, Seasonally Adjusted", xlab = "", ylab = "Burglaries")
```

```{r plot.tomv_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(offense %in% c("VEH. THEFT")) %>%
  group_by(date) %>%
  summarize(offenses = n()) %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
  labs(
    x = "", 
    y = "Vehicle Thefts", 
    title = "Vehicle Thefts", 
    caption = spokane_caption
    ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.assault_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(offense %in% c("ASSAULT")) %>%
  group_by(date) %>%
  summarize(offenses = n()) %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
  labs(
    x = "", 
    y = "Assaults", 
    title = "Assaults", 
    caption = spokane_caption
    ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r plot.rape_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(offense %in% c("RAPE")) %>%
  group_by(date) %>%
  summarize(offenses = n()) %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  geom_smooth(method = "auto", color = "blue", size = geom_line_size) +
  geom_smooth(method = "lm", color = "red", size = geom_line_size) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
  labs(
    x = "", 
    y = "Sexual Assaults", 
    title = "Sexual Assaults", 
    caption = spokane_caption
    ) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

# Police Districts

Spokane has a total of eight police districts. Those eight are divided evenly between two main groups, the North and South Police Service Areas. The Spokane River separates the two service areas, with Division St. and other major thoroughfares serving as boundaries. Districts P1 - P4 are north of the Spokane River. Districts P5 - P8 are south of the river. Some offenses in the CompStat reports are listed as "OTH", meaning "Other".

* P1: Northwest District
* P2: Central District
* P3: Neva-Wood District
* P4: Northeast District
* P5: South Central District
* P6: Southeast District
* P7: Southwest District
* P8: Downtown District
* OTH: Other District


```{r plot.ytd_offenses_by_district, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2022) %>%
  group_by(district) %>%
  summarize(count = n()) %>%
  ggplot(aes(reorder(district, -count), count, label = count)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "Police Districts", 
    title = "Offenses by Police District, 2022 YTD", 
    caption = spokane_caption_exclude_dist
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.total_offenses_by_district, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  group_by(district) %>%
  summarize(count = n()) %>%
  ggplot(aes(reorder(district, -count), count, label = count)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "Police Districts", 
    title = "Offenses by Police District, 2017 - 2022",
    caption = spokane_caption_exclude_dist
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.total_offenses_by_violence_district, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  count(district, violence) %>% 
  ggplot(aes(x = reorder(district, -n, sum), y = n, fill = violence, label = n)) +
  geom_col() +
  geom_text(position = position_stack(vjust = .5)) +
  scale_fill_manual("Offense Type", values = c("Non-Violent" = "cadetblue3", "Violent" = "coral1")) +
  theme(legend.position = c(0.925, 0.925), legend.background = element_rect(size=0.5, linetype="solid", colour ="black")) +
  labs(
    x = "Police Districts", 
    y = "Offenses", 
    title = "Offense Violence by District, 2017-2022", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

```{r plot.ytd_offenses_by_violence_district, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2022) %>%
  count(district, violence) %>%
  ggplot(aes(x = reorder(district, -n, sum), y = n, fill = violence, label = n)) +
  geom_col() +
  geom_text(position = position_stack(vjust = .5)) +
  scale_fill_manual("Offense Type", values = c("Non-Violent" = "cadetblue3", "Violent" = "coral1")) +
  theme(legend.position = c(0.925, 0.925), legend.background = element_rect(size=0.5, linetype="solid", colour ="black")) +
  labs(
    x = "Police Districts", 
    y = "Offenses", 
    title = "Offense Violence by District, 2022 YTD", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
    )
```

## North, South Districts

```{r plot.boxplot_p1_p4_avg_daily_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(!district %in% c(NA, "OTH", "P5", "P6", "P7", "P8")) %>%
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Offenses", 
    title = "North Policing Districts, Daily Offenses by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  ) +
  facet_grid(rows = vars(district))
```

```{r plot.boxplot_p5_p8_avg_daily_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(!district %in% c(NA, "OTH", "P1", "P2", "P3", "P4")) %>%
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Offenses", 
    title = "South Policing Districts, Daily Offenses by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  ) +
  facet_grid(rows = vars(district))
```

```{r plot.district_nas_over_time, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(is.na(district)) %>%
  group_by(date) %>%
  summarize(offenses = n()) %>%
  ggplot(aes(date, offenses)) + 
  geom_point(size = point_size, alpha = point_alpha) +
  labs(
    x = "", 
    y = "Offenses Missing District", 
    title = "Offenses Without a Police District", 
    caption = spokane_caption
    ) 
```


```{r plot.ytd_offenses_by_type, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2022) %>%
  count(offense) %>%
  ggplot(aes(reorder(offense, -n), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1)) +
  labs(
    x = "", 
    title = "Offense Types, 2022 YTD", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.total_offenses_by_type, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  count(offense) %>%
  ggplot(aes(reorder(offense, -n), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1)) +
  labs(
    x = "", 
    title = "Offense Types, 2017 - 2022", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.ytd_offenses_by_weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2022) %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Offenses by Day, 2022 YTD", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.2021_offenses_by_weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2021) %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Offenses by Day, 2021", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.2020_offenses_by_weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2020) %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Offenses by Day, 2020", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.2019_offenses_by_weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2019) %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Offenses by Day, 2019", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.2018_offenses_by_weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(year == 2018) %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Offenses by Day, 2018", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.Spokane_Offenses_By_Weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Spokane Offenses by Day 2017 - 2022", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.Bellevue_Offenses_By_Weekday, fig.pos="H", fig.width=10, fig.height=7.5}
df.bellevue_crimes %>%
  count(dow) %>%
  ggplot(aes(fct_relevel(dow,"Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "", 
    title = "Bellevue Offenses by Day", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.offenses_by_lunar_phase, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  count(lunar_phase) %>%
  ggplot(aes(fct_relevel(lunar_phase,"Full","Waning Gibb.","Last quarter","Waning Cres.","New","Waxing Cres.","First quarter","Waxing Gibb."), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1)) +
  labs(
    x = "Lunar Phases", 
    title = "Offenses by Lunar Phase", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

```{r plot.violent_offenses_by_lunar_phase, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(violence == "Violent") %>%
  count(lunar_phase) %>%
  ggplot(aes(fct_relevel(lunar_phase,"Full","Waning Gibb.","Last quarter","Waning Cres.","New","Waxing Cres.","First quarter","Waxing Gibb."), n, label = n)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25, hjust=1)) +
  labs(
    x = "Lunar Phase", 
    title = "Violent Offenses by Lunar Phase", 
    caption = spokane_caption
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```

# Seasonal

## Assault

```{r plot.assault_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(offense == "ASSAULT") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date
freq <- 365 # Time series frequency

### All data time series ###
ts.crimes <- ts(model_data, start = c(2017,start_date), end = c(2021,end_date), frequency = freq)

# Decompose time series
ts.crime_stl <- stl(ts.crimes[,2],"periodic") # Decompose time series
plot(ts.crime_stl, main = "Assault Time Series Decomposition")
```

## Arson

```{r plot.arson_by_district, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>%
  filter(offense == "ARSON") %>%
  group_by(district) %>%
  summarize(count = n()) %>% 
  ggplot(aes(reorder(district, -count), count, label = count)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  labs(
    x = "Police Districts", 
    title = "Arson by Police District", 
    caption = spokane_caption_exclude_dist
    ) +
  scale_y_continuous("Offenses", labels = comma, expand = c(.1, .1))
```


## Burglary

```{r plot.boxplot_avg_daily_burglaries_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>% 
  filter(offense == "BURGLARY") %>%
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Burglaries", 
    title = "Daily Burglaries by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  )
```

```{r plot.burglary_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(offense == "BURGLARY") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date
freq <- 365 # Time series frequency

### All data time series ###
ts.crimes <- ts(model_data, start = c(2017,start_date), end = c(2021,end_date), frequency = freq)

# Decompose time series
ts.crime_stl <- stl(ts.crimes[,2],"periodic") # Decompose time series
plot(ts.crime_stl, main = "Burglary Time Series Decomposition")
```

## Theft

```{r plot.boxplot_avg_daily_thefts_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>% 
  filter(offense == "THEFT") %>%
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Thefts", 
    title = "Daily Thefts by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  )
```

```{r plot.theft_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(offense == "THEFT") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date
freq <- 365 # Time series frequency

### All data time series ###
ts.crimes <- ts(model_data, start = c(2017,start_date), end = c(2021,end_date), frequency = freq)

# Decompose, plot time series
ts.crime_stl <- stl(ts.crimes[,2], s.window = "periodic", robust = TRUE)
plot(ts.crime_stl, main = "Theft Time Series Decomposition")
```

```{r plot.theft_time_series_seasonal_adj, fig.pos="H", fig.width=10, fig.height=7.5}
ts.seasonal   <- ts.crime_stl$time.series[,1]
ts.trend	   <- ts.crime_stl$time.series[,2]
ts.remainder  <- ts.crime_stl$time.series[,3]
 
plot(ts.trend + ts.remainder,
main="Thefts over Time, Seasonally Adjusted",
ylab="Thefts")
```

## Robbery

```{r plot.robbery_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(offense == "ROBBERY") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date
freq <- 365 # Time series frequency

### All data time series ###
ts.crimes <- ts(model_data, start = c(2017,start_date), end = c(2021,end_date), frequency = freq)

# Decompose time series
ts.crime_stl <- stl(ts.crimes[,2],"periodic") # Decompose time series
plot(ts.crime_stl, main = "Robbery Time Series Decomposition")
```

## Motor Vehicle Theft

```{r plot.boxplot_avg_daily_vehtheft_by_month, fig.pos="H", fig.width=10, fig.height=7.5}
df.crimes %>% 
  filter(offense == "VEH. THEFT") %>%
  group_by(day=floor_date(date, "day")) %>%
  mutate(day.count = n()) %>%
  group_by(month=floor_date(date, "month")) %>% 
  ggplot(aes(x = month)) +
  geom_boxplot(aes(y = day.count, group = month)) +
  labs(
    x = "", 
    y = "Vehicle Thefts", 
    title = "Daily Vehicle Thefts by Month", 
    caption = paste(spokane_data_source_str, cc_by_sa, sep = "\n")
  )
```

```{r plot.veh_theft_time_series_decomposition, fig.pos="H", fig.width=10, fig.height=7.5}
# Total Offenses by date
model_data <- df.crimes %>%
  filter(offense == "VEH. THEFT") %>%
  group_by(date) %>%
  summarize(offenses = n())

# Calculate start, end numerical dates
start_date <- as.numeric(as.Date("2017-08-01") - as.Date("2017-01-01")) # First 2017 observed date
end_date <- as.numeric(as.Date("2021-01-30") - as.Date("2021-01-01")) # Last 2021 observed date
freq <- 365 # Time series frequency

### All data time series ###
ts.crimes <- ts(model_data, start = c(2017,start_date), end = c(2021,end_date), frequency = freq)

# Decompose, plot time series
ts.crime_stl <- stl(ts.crimes[,2], s.window = "periodic", robust = TRUE)
plot(ts.crime_stl, main = "Motor Vehicle Theft Time Series Decomposition")
```

```{r plot.veh_theft_time_series_seasonal_adj, fig.pos="H", fig.width=10, fig.height=7.5}
ts.seasonal   <- ts.crime_stl$time.series[,1]
ts.trend	   <- ts.crime_stl$time.series[,2]
ts.remainder  <- ts.crime_stl$time.series[,3]
 
plot(ts.trend + ts.remainder,
main="Vehicle Thefts over Time, Seasonally Adjusted",
ylab="Thefts")
```